---
title: "ESM 206 Lab 5"
author: "Pat Byrne"
date: "10/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Lab 5 objectives

- Overthrow the ruling class
- Seize the means of productions
- Learn to pase dates with lubrication
- Learn to count
- Learn to uncount
- (s)t(d)-tests because safe sex is important
- Create a hotmap


#### Setup

```{r, include = F}
# Attach packages
library(tidyverse)
library(here)
library(janitor)

# Read in data, clean it up a lil bit
lobster.abundance <- read_csv(here('data','lobster_abundance.csv'),
                              na = '-99999') %>% 
  clean_names()
```
Every observation isn't its own row! There are counts for each combination date, site, transect, replicate, and size. Aka, each observation is not represented

We'll use uncount() to convert our lobster data from frequency format to case format
```{r}
lobster.tidy <- lobster.abundance %>% 
  uncount(lobster_count)
```


#### Exploratory data visualization

We're only going to consider site as our variable of interest
```{r}
#Jitter plot
ggplot(data = lobster.tidy, aes(x = site, y = size_mm)) +
  geom_jitter(aes(color = site),
              alpha = 0.5,
              width = 0.2)

# Histogram
ggplot(lobster.tidy, aes(x = size_mm)) +
  geom_histogram(aes(fill = site)) +
  facet_wrap(~site, scales = 'free')

# QQ plot
ggplot(lobster.tidy, aes(sample = size_mm)) +
  geom_qq() + 
  facet_wrap(~site)
```


#### Convert date column to class 'Date'

We'll use lubrication to convert the format and easily parse month and year
```{r}

lobster.date <- lobster.tidy %>% 
  mutate(
    date_new = lubridate::mdy(date)
  )

```

Parse year and month
```{r}

lobster.parse <- lobster.date %>% 
  mutate(
    obs_month = lubridate::month(date_new, 
                                 label = T),
    obs_year = lubridate::year(date_new)
  )

```

Let's find counts of observed lobsters based on different hierarchical groupings

First let's count lobsters by year and month
```{r}

lobster.ym <- lobster.parse %>% 
  count(obs_year, obs_month)

lobster.y <- lobster.parse %>% 
  count(obs_year)

lobster.s <- lobster.parse %>% 
  count(site)

```

If we want to create a summary table that contains statistics OTHER than counts by group, it's easier to use group_by() + n()
```{r}

lobster_summary <- lobster.parse %>% 
  group_by(site) %>% 
  summarize(
    mean_size = mean(size_mm, na.rm = T),
    sd_size = sd(size_mm, na.rm = T),
    lobster_number = n()
  )

```

#### Find confidence interval

Use t.test() to find confidence intervals for one sample and perform t-tests to compare means of two samples

